\documentclass[12pt,letterpaper]{article}
\usepackage{graphicx,textcomp}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{fullpage}
\usepackage{color}
\usepackage[reqno]{amsmath}
\usepackage{amsthm}
\usepackage{fancyvrb}
\usepackage{amssymb,enumerate}
\usepackage[all]{xy}
\usepackage{endnotes}
\usepackage{lscape}
\newtheorem{com}{Comment}
\usepackage{float}
\usepackage{hyperref}
\newtheorem{lem} {Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{obs}{Observation}
\usepackage[compact]{titlesec}
\usepackage{dcolumn}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usepackage{multirow}
\usepackage{xcolor}
\newcolumntype{.}{D{.}{.}{-1}}
\newcolumntype{d}[1]{D{.}{.}{#1}}
\definecolor{light-gray}{gray}{0.65}
\usepackage{url}
\usepackage{listings}
\usepackage{color}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
	backgroundcolor=\color{backcolour},   
	commentstyle=\color{codegreen},
	keywordstyle=\color{magenta},
	numberstyle=\tiny\color{codegray},
	stringstyle=\color{codepurple},
	basicstyle=\footnotesize,
	breakatwhitespace=false,         
	breaklines=true,                 
	captionpos=b,                    
	keepspaces=true,                 
	numbers=left,                    
	numbersep=5pt,                  
	showspaces=false,                
	showstringspaces=false,
	showtabs=false,                  
	tabsize=2
}
\lstset{style=mystyle}
\newcommand{\Sref}[1]{Section~\ref{#1}}
\newtheorem{hyp}{Hypothesis}

\title{Problem Set 1}
\author{Yana Konshyna}


\begin{document}
	\maketitle
	\section*{Instructions}
	\begin{itemize}
	\item Please show your work! You may lose points by simply writing in the answer. If the problem requires you to execute commands in \texttt{R}, please include the code you used to get your answers. Please also include the \texttt{.R} file that contains your code. If you are not sure if work needs to be shown for a particular problem, please ask.
\item Your homework should be submitted electronically on GitHub in \texttt{.pdf} form.
\item This problem set is due before 23:59 on Sunday February 11, 2024. No late assignments will be accepted.
	\end{itemize}

	\vspace{.25cm}
\section*{Question 1} 
\vspace{.25cm}
\noindent The Kolmogorov-Smirnov test uses cumulative distribution statistics test the similarity of the empirical distribution of some observed data and a specified PDF, and serves as a goodness of fit test. The test statistic is created by:

$$D = \max_{i=1:n} \Big\{ \frac{i}{n}  - F_{(i)}, F_{(i)} - \frac{i-1}{n} \Big\}$$

\noindent where $F$ is the theoretical cumulative distribution of the distribution being tested and $F_{(i)}$ is the $i$th ordered value. Intuitively, the statistic takes the largest absolute difference between the two distribution functions across all $x$ values. Large values indicate dissimilarity and the rejection of the hypothesis that the empirical distribution matches the queried theoretical distribution. The p-value is calculated from the Kolmogorov-
Smirnoff CDF:

$$p(D \leq d)= \frac{\sqrt {2\pi}}{d} \sum _{k=1}^{\infty }e^{-(2k-1)^{2}\pi ^{2}/(8d^{2})}$$


\noindent which generally requires approximation methods (see \href{https://core.ac.uk/download/pdf/25787785.pdf}{Marsaglia, Tsang, and Wang 2003}). This so-called non-parametric test (this label comes from the fact that the distribution of the test statistic does not depend on the distribution of the data being tested) performs poorly in small samples, but works well in a simulation environment. Write an \texttt{R} function that implements this test where the reference distribution is normal. Using \texttt{R} generate 1,000 Cauchy random variables (\texttt{rcauchy(1000, location = 0, scale = 1)}) and perform the test (remember, use the same seed, something like \texttt{set.seed(123)}, whenever you're generating your own data).\\
	
	
\noindent As a hint, you can create the empirical distribution and theoretical CDF using this code:

\begin{lstlisting}[language=R]
	# create empirical distribution of observed data
	ECDF <- ecdf(data)
	empiricalCDF <- ECDF(data)
	# generate test statistic
	D <- max(abs(empiricalCDF - pnorm(data))) 
	\end{lstlisting}

\vspace{.5cm}
\noindent 1.Creating data:
\lstinputlisting[language=R, firstline=40,lastline=42]{PS1_answers_YK_23359606.R} 

\vspace{.5cm}
\noindent 2.Creating function \texttt{ksTestFunction}, which accept our data and then calculate test-statistics and p-value:
\lstinputlisting[language=R, firstline=45,lastline=61]{PS1_answers_YK_23359606.R} 


\vspace{.5cm}
\noindent \textbf{Output: }
\begin{lstlisting}
$D
[1] 0.1347281

$P_value
[1] 5.652523e-29
\end{lstlisting}

\vspace{.5cm}
\noindent 3. Checking the results by using ks.test() built-in function:
\lstinputlisting[language=R, firstline=64,lastline=64]{PS1_answers_YK_23359606.R} 

\vspace{.5cm}
\noindent \textbf{Output: }
\begin{lstlisting}
Asymptotic one-sample Kolmogorov-Smirnov 

testdata:  cauchyData
D = 0.13573, p-value = 2.22e-16
alternative hypothesis: two-sided
\end{lstlisting}

\pagebreak

\section*{Question 2}
\noindent Estimate an OLS regression in \texttt{R} that uses the Newton-Raphson algorithm (specifically \texttt{BFGS}, which is a quasi-Newton method), and show that you get the equivalent results to using \texttt{lm}. Use the code below to create your data.
\vspace{.5cm}
\lstinputlisting[language=R, firstline=72,lastline=74]{PS1_answers_YK_23359606.R} 

\vspace{.5cm}
\noindent 1. Creating data:
\lstinputlisting[language=R, firstline=72,lastline=74]{PS1_answers_YK_23359606.R} 

\vspace{.5cm}
\noindent The probability density function (PDF) for normal distribution is

\begin{equation} 
f(x)= \frac{1}{\sigma\sqrt {2\pi}} e^{-\frac{1}{2} (\frac{y_i-x_i\beta}{\sigma})^2}
\end{equation}

\vspace{.5cm}
\noindent To find the parameters I use log of likelihood function:

\begin{equation} 
	\log(\mathcal{L}) = \log(f(x))= -\frac{1}{2} n \log(2\pi) - \frac{1}{2} n \log(\sigma^2) - \frac{1}{2\sigma^2} \sum_{i=1}^{n} (y_i - x_i\beta)^2 
\end{equation}

\vspace{.5cm}
\noindent 2. Creating log-likelihood function using formula (2):
\lstinputlisting[language=R, firstline=77,lastline=90]{PS1_answers_YK_23359606.R} 

\vspace{.5cm}
\noindent 3. Doing optimization using \texttt{optim} function by using BFGS method. This function will adjust the \texttt{par} vector to minimize the negative log-likelihood, finding the best-fitting linear model to the data given the assumption of normally distributed errors:
\lstinputlisting[language=R, firstline=93,lastline=94]{PS1_answers_YK_23359606.R} 

\vspace{.5cm}
\noindent 4. Display optimization results:
\lstinputlisting[language=R, firstline=97,lastline=97]{PS1_answers_YK_23359606.R} 

\vspace{.5cm}
\noindent \textbf{Output: }
\begin{lstlisting}
0.14 2.73
 \end{lstlisting}

\vspace{.5cm}
\noindent 5. Checking the results by performing OLS regression using lm function
\lstinputlisting[language=R, firstline=100,lastline=101]{PS1_answers_YK_23359606.R} 

\begin{table}[H] \centering   \caption{OLS regression}   \label{} \begin{tabular}{@{\extracolsep{5pt}}lc} \\[-1.8ex]\hline \hline \\[-1.8ex]  & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ \cline{2-2} \\[-1.8ex] & y \\ \hline \\[-1.8ex]  x & 2.727$^{***}$ \\   & (0.042) \\   & \\  Constant & 0.139 \\   & (0.253) \\   & \\ \hline \\[-1.8ex] Observations & 200 \\ R$^{2}$ & 0.956 \\ Adjusted R$^{2}$ & 0.956 \\ Residual Std. Error & 1.447 (df = 198) \\ F Statistic & 4,298.687$^{***}$ (df = 1; 198) \\ \hline \hline \\[-1.8ex] \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ \end{tabular} \end{table} 

\end{document}
